---
title: "<center>Assignment8<center>"
author: "<center>Guandi Xu<center>"
date: "<center>10/30/2020<center>"
output: pdf_document
---


## 5.3.3 Rejection Sampling

#### 1. Show the result

As given, the O-U process is:
$$
dr(t) = \alpha (b-r(t))dt + \sigma dW(t)
$$
According to the Ito formula:
$$
dV(t) = \frac{\partial V}{\partial t}dt + \frac{\partial V}{\partial r(t)}dr(t) + \frac{1}{2} \frac{\partial ^2 V}{\partial r(t)^2} dr(t)^2
$$
Let $V(t) = \exp(\alpha t)(r(t) - b)$,
then we have 
$\frac{\partial V}{\partial t} = \alpha \exp(\alpha t)(r(t) - b)$, 
$\frac{\partial V}{\partial r(t)} = \exp(\alpha t)$, 
$\frac{\partial^2 V}{\partial r(t)^2} = 0$, 
Therefore, we can get,
$$
dV(t) = \alpha \exp(\alpha t)(r(t)-b)dt + \exp(\alpha t)dr(t)
$$
$$
dV(t) = \sigma \exp(\alpha t) dW(t)
$$
Then we can take the integral from $t$ to $t + \Delta$
$$
\int_t^{t+\Delta} dV(u) = \int_t^{t+\Delta} \sigma \exp(\alpha u)dW(u)
$$
From the left hand side, we can get
$$
\int_t^{t+\Delta} dV(u) = \exp[\alpha(t+\Delta)] (r(t + \Delta) -b) - \exp(\alpha t)(r(t) - b)
$$
Then, we can try to take the $r(t+\Delta)$ out and calculate the result:
$$
r(t + \Delta) - b = \exp(-\alpha \Delta)(r(t) - b) + \int_t^{t+\Delta} \sigma \exp[-\alpha (t + \Delta -u)] dW(u) \\
r(t + \Delta) = b + \exp(-\alpha \Delta)(r(t) - b) + \int_t^{t+\Delta} \sigma \exp[-\alpha (t + \Delta -u)] dW(u)
$$
According to the property of the Brownian Motion, we know that
$$
\int_t^{t+\Delta} \sigma \exp[-\alpha (t+\Delta-u)] dW(u) \sim N(0, \int_t^{t+\Delta} \sigma^2 \exp(2*[-\alpha (t+\Delta-u)]) dW(u) )
$$
So, we can get that 
$$
r(t+\Delta)=b+e^{-\alpha\Delta}(r(t)-b)+N[0,\int_t^{t+\Delta}\sigma^2 e^{-2\alpha(t+\Delta-u)}du] \\
r(t+\Delta)=e^{-\alpha\Delta}r(t)+b(1-e^{-\alpha\Delta})+\frac{\sigma}{\sqrt{2\alpha}}\sqrt{1-e^{-2\alpha\Delta}}Z
$$

#### 2. Transition Distribution

```{r}
ou <- function(alpha,sigma, b){
  r <- rep(0,500)
  r[1] <- 1
  t <- 500
  delta <- 1/500
for (i in 2:500){
  r[i] <- exp(-alpha)*r[i-1]+b*(1-exp(-alpha))+
    (sigma/sqrt(2*alpha))*sqrt(1-exp(-2*alpha))*rnorm(1)
}
  return(r)
}
##For example
plot(ou(0.1,0.1,-5))
##Aggregate
par(mfrow=c(3,2))
plot(ou(0.1,0.1,-5));plot(ou(0.1,0.2,-5));plot(ou(0.1,0.5,-5))
plot(ou(1,0.1,-5));plot(ou(1,0.2,-5));plot(ou(1,0.5,-5))
par(mfrow=c(3,2))
plot(ou(5,0.1,-5));plot(ou(5,0.2,-5));plot(ou(5,0.5,-5))
plot(ou(0.1,0.1,5));plot(ou(0.1,0.2,5));plot(ou(0.1,0.5,5))
par(mfrow=c(3,2))
plot(ou(1,0.1,5));plot(ou(1,0.2,5));plot(ou(1,0.5,5))
plot(ou(5,0.1,5));plot(ou(5,0.2,5));plot(ou(5,0.5,5))
```
Here we can see that r(t) increases towards 5 and decreases towards -5. which means b determines the target position, $\alpha$ determines the speed going to b, and $\sigma$ determines the variation.

#### 3. Use Euler-Maruyama Method

Euler–Maruyama method is:
$$
Y_{n+1}=Y_n + a(Y_n)\Delta t+b(Y_n)\Delta W_n
$$
In our O-U process, T=1, number of steps is $n=\frac{1}{\delta}$, $a(Y_n)=\alpha (b-r(t))$, and $b(Y_n)=\sigma$.
We can set $r(0)=r(t_0)=0$, our goal is sampling $r(1)=r(t_{n+1})$, n={1,2,10,100}.
```{r}
euler <- function(delta,alpha,sigma,b){
  n <- 1/delta
  r <- rep(0,n+1)
  r[1] <- 1
  for (i in 2:(n+1)){
    r[i] <- r[i-1]+ alpha * (b-r[i-1]) * delta + sigma * rnorm(1,0,delta)
  }
  r[n+1]
}
delta <- c(1,0.5,0.1,0.01)
x <- matrix(0,4,1000)
for (j in 1:4) {
  for (k in 1:1000) {
    ##We choose b=1 so that it is martingale
    ##r(1) should have mean=1
    x[j,k] <- euler(delta[j],1,0.1,1)
  }
}
plot(density(x[1,]),main = "delta=1")
lines(seq(0.5,1.5,0.01),dnorm(seq(0.5,1.5,0.01),1,0.1),col="red")
legend("topleft", c("True","Estimated"), col = c("red","black"),cex = 1,lwd = 1)
plot(density(x[2,]),main = "delta=0.5")
lines(seq(0.5,1.5,0.01),dnorm(seq(0.5,1.5,0.01),1,0.1),col="red")
legend("topleft", c("True","Estimated"), col = c("red","black"),cex = 1,lwd = 1)
plot(density(x[3,]),main = "delta=0.1")
lines(seq(0.5,1.5,0.01),dnorm(seq(0.5,1.5,0.01),1,0.1),col="red")
legend("topleft", c("True","Estimated"), col = c("red","black"),cex = 1,lwd = 1)
plot(density(x[4,]),main = "delta=0.01")
lines(seq(0.5,1.5,0.01),dnorm(seq(0.5,1.5,0.01),1,0.1),col="red")
legend("topleft", c("True","Estimated"), col = c("red","black"),cex = 1,lwd = 1)

```

From the graphs showed above, we can clearly see that when $a(Y_n)=\alpha (b-r(t))=0$, this SDE became martingale. And more steps we take, higher density it gets, which means more accruate.



